---
title: "Final Project Proposal"
author: "Daniel Dittenhafer & Justin Hink"
date: "March 1, 2016"
documentclass: article 
geometry: margin=0.75in
output: 
  pdf_document:
    number_sections: yes
subtitle: "DATA 621: Business Analytics and Data Mining"
---
```{r, echo=FALSE, message=FALSE}
library(knitr)
library(knitcitations)
library(RefManageR)

cleanbib()

cite_options(style="markdown")

bibBirthData <- bibentry(bibtype="Misc",
                         author=person(family="HHS"),
                         publisher="United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics",
                         title="Natality public-use data 2007-2014 on CDC WONDER Online Database",
                         year=2016,
                         month="February",
                         url="http://wonder.cdc.gov/natality-current.html")

bibBirthData06 <- bibentry(bibtype="Misc",
                         author=person(family="HHS"),
                         publisher="United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics",
                         title="Natality public-use data 2003-2006 on CDC WONDER Online Database",
                         year=2009,
                         month="March",
                         url="http://wonder.cdc.gov/natality-v2006.html")

bibSpMorg21 <- bibentry(bibtype="Misc",
                 author=personList(person(family="Morgan", given="S.", middle="Philip"), 
                                   person(family="Miles", given="G.", middle="Taylor")),
                 journal="Annual Review of Sociology",
                 title="Low Fertility at the Turn of the Twenty-First Century",
                 year=2006,
                 month="August",
                 url="http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2849172/")
```

# Data Set

The data sets we plan to use originate from United States federal government agencies. The core data comes from the U.S. Department of Health & Human Services Centers for Disease Control and Prevention  `r citep(bibBirthData)`. Related population and economic data sets will come from the U.S. Census Bureau and the U.S. Bureau of Labor Statistics.


## Natality Data

Separate data sets the years 2003-2006, and 2007-2014 were downloaded from the CDC website `r citep(bibBirthData06)`, `r citep(bibBirthData)`.

```{r, echo=FALSE}
dfNatality <- c("Year", "Month",	"Age of Mother",	"Marital Status",	"Education",	"Births")
kable(dfNatality)
```

##  Census Data

**TBD: Data set descriptions**


## Unemployment Data

**TBD: Data set descriptions**


# Recent Journal Papers

**TBD**

## Low Fertility at the Turn of the Twenty-First Century

The Annual Review of Sociology included a paper by S. Philip Morgan and Miles G. Taylor regarding recent  fertility trends, specifically a shift to lower birth rates as compared to the second half of the twentieth century `r citep(bibSpMorg21)`.

# Problem Statement & Research Questions

**TBD**

# Statement of Objectives

**TBD**


# References

```{r, results='asis', echo=FALSE}
BibOptions(style="html", bib.style="authortitle")
bibliography()
```

